working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect

### 2019-03-06 Manuscript_draft/Manuscript_draft.Rmd
edits

**Hand-pollination**

@@

The traits measured for each species were pollen per anther, number of ovules and stigma, style, ovary width and length. For the stigma, the stigmatic area was also measured and moreover the stigmas were divided in wet/dry type. All the morphometrical measurements were performed with a stereophotomicroscope. Pollen was counted for 20 anthers of each species with 4 replicates per sample with an hemocytometer. Previously, anthers were squashed on a known solution with the pippete tip and homogeneize with a vortex for 30 seconds. Ovule number was counted with the help of a stereomicroscope and a small grid over a petri dish from 15 randomly selected flowers. Fruits per number of flowers treated were counted for Solanaceae species and for the rest, the number of seeds produced per average number of ovules of each species. Levels of self incompatibility were estimated by dividing the fruit set of hand self pollination by hand cross pollination @lloyd1992.

@@

In order to test the relative effect of traits on seed production with foreign pollen we performed Mantel test between the assymetrical matrix of heterospecific pollen effect (10 by 10 matrix) with the different distance matrices of traits (Euclidean distances). Heterospecific pollen effect was obtained through the subtraction of seed production by hand cross pollination minus seed production of the different heterospecific pollen treatments. To find a model with the best explanatory traits we used the function **bioenv** from R. We also conducted Mantel test between the matrix of heterospecific pollen effect and the distance matrix from all the traits. Moreover, we explored also the correlation between traits and heterospecific pollen effect through generalized mixed models where the response variable was heterospecific pollen effect and the explanatory variable the different traits. In addition, we tested the correlation between the total amount of pollen deposited on the stigma with heterospecific pollinations and the stigma size through Pearson's correlation.

@@

Results of hand cross-pollination, self hand-pollination, natural selfing and apomixis are presented in **Figure 1** (see appendix 1 for table with values). Heterospecific pollen reduced seet set significatively with the 50-50% heterospecific pollen treatments for 65% of the pairwise interactions p<0.05. Moreover, effect sizes differ across species, see **Figure 2** where species like *S. lycopersicum* had a large effect size and other's like *E. sativa* a small or null effect size.

Across families we found a very similar effect but when species where look at species level they respond differently even within the same family [rephrase and maybe test statistically?](), for instance for two species of the Brassicaceae family *Brassica oleracea* and *Eruca versicaria* we found very contrasting effects of foreign pollen where for the first one, all donors reduce seed set significatively and for the second, just two species did out of nine. The 100% foreign pollen treatments barely produced seeds or fruits [and just for *Sinapis alba* we did not find significant differences between the hand cross pollination and one treatment with pollen from a confamilial- IB Unclear](). Solanaceae species with berry fruit type developed small fruits or even normal fruits in some cases [under which treatment](). *S. lycopersicum* seems to produced small fruits (35% of the treatments) independently of pollen and pollen donor due to also apomictic treatments did, never normal size. *C. annuum* produced some fruits (9%) of both small and normal size and finally *S. melongena* produced seedless normal fruits with just confamilial pollen (3%), for both species seems that fruit formation was induced by pollen on the stigma because of lack of fruit production with treatments that tested for apomixis.[clarify this descriptive statistics part- Also a figure with a summary of the treatments effect would be cool, or at least in the appendices]()

\newpage
```{r, echo=FALSE,fig.align='center',message=FALSE, warning=FALSE, cache=FALSE,fig.height=4,fig.width=7, fig.cap="Barplot with the different treatments that provide information of the reproductive biology of the ten species. The y axis is the proportion of ovules converted to seed in percentage. The different treatments (N=10) which are presented in the legend are, hand cross pollination, hand self pollination, natural selfing and apomixis. More information about these treatments can be found in Methods and Appendices."}

@@

**Table S1**. Perecentage of seeds produced per ovule for the ten species used in the experiment. The treatments presented are hand cross pollination, hand self pollination, natural selfing and apomixis (emasculated flowers).
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
library(kableExtra)
load("table_treatments.RData")
Species <- c("Brassica oleracea", "Brassica rapa", "Eruca versicaria", "Sinapis alba", "Ipomoea aquatica", "Ipomoea purpurea", "Capsicum annuum", "Petunia integrifolia", "Solanum lycopersicum", "Solanum melongena")
Cross <- cross
Self <- self
Natural_selfing <- nat_self
Apomixis <- apo
table<- data.frame(Species, Cross, Self, Natural_selfing,Apomixis)
kable(table, format="markdown")
```

>>>>

**Hand pollination**

@@

The traits measured for each species were pollen per anther, number of ovules and stigma, style, ovary width and length. For the stigma, the stigmatic area was also measured and moreover the stigmas were divided in wet/dry type. All the morphometrical measurements were performed with a stereophotomicroscope. Pollen was counted for 20 anthers of each species with 4 replicates per sample with an hemocytometer. Previously, anthers were squashed on a known solution with the pippete tip and homogeneize with a vortex for 30 seconds. Ovule number was counted with the help of a stereomicroscope and a small grid over a petri dish from 15 randomly selected flowers. Fruits per number of flowers treated were counted for just Solanaceae species, we counted for all the species the number of seeds produced per average number of ovules. Levels of self-incompatibility were estimated by dividing the fruit set of hand self-pollination by hand cross pollination @lloyd1992.

@@

In order to test the relative effect of traits on seed production with foreign pollen we performed Mantel test between the assymetrical matrix of heterospecific pollen effect (10 by 10 matrix) with the different distance matrices of traits (Euclidean distances). Heterospecific pollen effect was obtained through the subtraction of seed production by hand cross-pollination minus seed production of the different heterospecific pollen treatments. To find a model with the best explanatory traits we used the function **bioenv** from R. We also conducted Mantel test between the matrix of heterospecific pollen effect and the distance matrix from all the traits. Moreover, we explored also the correlation between traits and heterospecific pollen effect through generalized mixed models where the response variable was heterospecific pollen effect and the explanatory variable the different traits. In addition, we tested the correlation between the total amount of pollen deposited on the stigma with heterospecific pollinations and the stigma size through Pearson's correlation.

@@

Results of hand cross-pollination, hand self-pollination, natural selfing and apomixis are presented in **Figure 1** (see appendix 1 for table with values). Heterospecific pollen reduced seet set significatively with the 50-50% heterospecific pollen treatments for 65% of the pairwise interactions p<0.05. Moreover, effect sizes differ across species, see **Figure 2** where species like *S. lycopersicum* had a large effect size and other's like *E. sativa* a small or null effect size.

Across families we found a very similar effect but when species where look at species level they respond differently even within the same family [rephrase and maybe test statistically?](), for instance for two species of the Brassicaceae family *Brassica oleracea* and *Eruca versicaria* we found very contrasting effects of foreign pollen where for the first one, all donors reduce seed set significatively and for the second, just two species did out of nine. The 100% foreign pollen treatments barely produced seeds or fruits [and just for *Sinapis alba* we did not find significant differences between the hand cross-pollination and one treatment with pollen from a confamilial- IB Unclear](). Solanaceae species with berry fruit type developed small fruits or even normal fruits in some cases [under which treatment](). *S. lycopersicum* seems to produced small fruits (35% of the treatments) independently of pollen and pollen donor due to also apomictic treatments did, never normal size. *C. annuum* produced some fruits (9%) of both small and normal size and finally *S. melongena* produced seedless normal fruits with just confamilial pollen (3%), for both species seems that fruit formation was induced by pollen on the stigma because of lack of fruit production with treatments that tested for apomixis.[clarify this descriptive statistics part- Also a figure with a summary of the treatments effect would be cool, or at least in the appendices]()

\newpage
```{r, echo=FALSE,fig.align='center',message=FALSE, warning=FALSE, cache=FALSE,fig.height=4,fig.width=7, fig.cap="Barplot with the different treatments that provide information of the reproductive biology of the ten species. The y axis is the proportion of ovules converted to seed in percentage. The different treatments (N=10) which are presented in the legend are, hand cross-pollination, hand self-pollination, natural selfing and apomixis. More information about these treatments can be found in Methods and Appendices."}

@@

**Table S1**. Perecentage of seeds produced per ovule for the ten species used in the experiment. The treatments presented are hand cross-pollination, hand self-pollination, natural selfing and apomixis (emasculated flowers).
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
library(kableExtra)
load("table_treatments.RData")
Species <- c("Brassica oleracea", "Brassica rapa", "Eruca versicaria", "Sinapis alba", "Ipomoea aquatica", "Ipomoea purpurea", "Capsicum annuum", "Petunia integrifolia", "Solanum lycopersicum", "Solanum melongena")
Cross <- cross
Self <- self
Natural_selfing <- nat_self
Apomixis <- apo
table<- data.frame(Species, Cross, Self, Natural_selfing,Apomixis)
kable(table, format="markdown")
```

### 2019-03-06 Manuscript_draft/Manuscript_draft.tex
edits

\textbf{Hand-pollination}

@@

The traits measured for each species were pollen per anther, number of
ovules and stigma, style, ovary width and length. For the stigma, the
stigmatic area was also measured and moreover the stigmas were divided
in wet/dry type. All the morphometrical measurements were performed with
a stereophotomicroscope. Pollen was counted for 20 anthers of each
species with 4 replicates per sample with an hemocytometer. Previously,
anthers were squashed on a known solution with the pippete tip and
homogeneize with a vortex for 30 seconds. Ovule number was counted with
the help of a stereomicroscope and a small grid over a petri dish from
15 randomly selected flowers. Fruits per number of flowers treated were
counted for Solanaceae species and for the rest, the number of seeds
produced per average number of ovules of each species. Levels of self
incompatibility were estimated by dividing the fruit set of hand self
pollination by hand cross pollination Lloyd and Schoen (1992).

@@

In order to test the relative effect of traits on seed production with
foreign pollen we performed Mantel test between the assymetrical matrix
of heterospecific pollen effect (10 by 10 matrix) with the different
distance matrices of traits (Euclidean distances). Heterospecific pollen
effect was obtained through the subtraction of seed production by hand
cross pollination minus seed production of the different heterospecific
pollen treatments. To find a model with the best explanatory traits we
used the function \textbf{bioenv} from R. We also conducted Mantel test
between the matrix of heterospecific pollen effect and the distance
matrix from all the traits. Moreover, we explored also the correlation
between traits and heterospecific pollen effect through generalized
mixed models where the response variable was heterospecific pollen
effect and the explanatory variable the different traits. In addition,
we tested the correlation between the total amount of pollen deposited
on the stigma with heterospecific pollinations and the stigma size
through Pearson's correlation.

@@

Results of hand cross-pollination, self hand-pollination, natural
selfing and apomixis are presented in \textbf{Figure 1} (see appendix 1
for table with values). Heterospecific pollen reduced seet set
significatively with the 50-50\% heterospecific pollen treatments for
65\% of the pairwise interactions p\textless{}0.05. Moreover, effect
sizes differ across species, see \textbf{Figure 2} where species like
\emph{S. lycopersicum} had a large effect size and other's like \emph{E.
sativa} a small or null effect size.

Across families we found a very similar effect but when species where
look at species level they respond differently even within the same
family \href{}{rephrase and maybe test statistically?}, for instance for
two species of the Brassicaceae family \emph{Brassica oleracea} and
\emph{Eruca versicaria} we found very contrasting effects of foreign
pollen where for the first one, all donors reduce seed set
significatively and for the second, just two species did out of nine.
The 100\% foreign pollen treatments barely produced seeds or fruits
\href{}{and just for \emph{Sinapis alba} we did not find significant
differences between the hand cross pollination and one treatment with
pollen from a confamilial- IB Unclear}. Solanaceae species with berry
fruit type developed small fruits or even normal fruits in some cases
\href{}{under which treatment}. \emph{S. lycopersicum} seems to produced
small fruits (35\% of the treatments) independently of pollen and pollen
donor due to also apomictic treatments did, never normal size. \emph{C.
annuum} produced some fruits (9\%) of both small and normal size and
finally \emph{S. melongena} produced seedless normal fruits with just
confamilial pollen (3\%), for both species seems that fruit formation
was induced by pollen on the stigma because of lack of fruit production
with treatments that tested for apomixis.\href{}{clarify this
descriptive statistics part- Also a figure with a summary of the
treatments effect would be cool, or at least in the appendices}

@@

\caption{Barplot with the different treatments that provide information of the reproductive biology of the ten species. The y axis is the proportion of ovules converted to seed in percentage. The different treatments (N=10) which are presented in the legend are, hand cross pollination, hand self pollination, natural selfing and apomixis. More information about these treatments can be found in Methods and Appendices.}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\textbf{Table S1}. Perecentage of seeds produced per ovule for the ten
species used in the experiment. The treatments presented are hand cross
pollination, hand self pollination, natural selfing and apomixis
(emasculated flowers).

>>>>

\textbf{Hand pollination}

@@

The traits measured for each species were pollen per anther, number of
ovules and stigma, style, ovary width and length. For the stigma, the
stigmatic area was also measured and moreover the stigmas were divided
in wet/dry type. All the morphometrical measurements were performed with
a stereophotomicroscope. Pollen was counted for 20 anthers of each
species with 4 replicates per sample with an hemocytometer. Previously,
anthers were squashed on a known solution with the pippete tip and
homogeneize with a vortex for 30 seconds. Ovule number was counted with
the help of a stereomicroscope and a small grid over a petri dish from
15 randomly selected flowers. Fruits per number of flowers treated were
counted for just Solanaceae species, we counted for all the species the
number of seeds produced per average number of ovules. Levels of
self-incompatibility were estimated by dividing the fruit set of hand
self-pollination by hand cross pollination Lloyd and Schoen (1992).

@@

In order to test the relative effect of traits on seed production with
foreign pollen we performed Mantel test between the assymetrical matrix
of heterospecific pollen effect (10 by 10 matrix) with the different
distance matrices of traits (Euclidean distances). Heterospecific pollen
effect was obtained through the subtraction of seed production by hand
cross-pollination minus seed production of the different heterospecific
pollen treatments. To find a model with the best explanatory traits we
used the function \textbf{bioenv} from R. We also conducted Mantel test
between the matrix of heterospecific pollen effect and the distance
matrix from all the traits. Moreover, we explored also the correlation
between traits and heterospecific pollen effect through generalized
mixed models where the response variable was heterospecific pollen
effect and the explanatory variable the different traits. In addition,
we tested the correlation between the total amount of pollen deposited
on the stigma with heterospecific pollinations and the stigma size
through Pearson's correlation.

@@

Results of hand cross-pollination, hand self-pollination, natural
selfing and apomixis are presented in \textbf{Figure 1} (see appendix 1
for table with values). Heterospecific pollen reduced seet set
significatively with the 50-50\% heterospecific pollen treatments for
65\% of the pairwise interactions p\textless{}0.05. Moreover, effect
sizes differ across species, see \textbf{Figure 2} where species like
\emph{S. lycopersicum} had a large effect size and other's like \emph{E.
sativa} a small or null effect size.

Across families we found a very similar effect but when species where
look at species level they respond differently even within the same
family \href{}{rephrase and maybe test statistically?}, for instance for
two species of the Brassicaceae family \emph{Brassica oleracea} and
\emph{Eruca versicaria} we found very contrasting effects of foreign
pollen where for the first one, all donors reduce seed set
significatively and for the second, just two species did out of nine.
The 100\% foreign pollen treatments barely produced seeds or fruits
\href{}{and just for \emph{Sinapis alba} we did not find significant
differences between the hand cross-pollination and one treatment with
pollen from a confamilial- IB Unclear}. Solanaceae species with berry
fruit type developed small fruits or even normal fruits in some cases
\href{}{under which treatment}. \emph{S. lycopersicum} seems to produced
small fruits (35\% of the treatments) independently of pollen and pollen
donor due to also apomictic treatments did, never normal size. \emph{C.
annuum} produced some fruits (9\%) of both small and normal size and
finally \emph{S. melongena} produced seedless normal fruits with just
confamilial pollen (3\%), for both species seems that fruit formation
was induced by pollen on the stigma because of lack of fruit production
with treatments that tested for apomixis.\href{}{clarify this
descriptive statistics part- Also a figure with a summary of the
treatments effect would be cool, or at least in the appendices}

@@

\caption{Barplot with the different treatments that provide information of the reproductive biology of the ten species. The y axis is the proportion of ovules converted to seed in percentage. The different treatments (N=10) which are presented in the legend are, hand cross-pollination, hand self-pollination, natural selfing and apomixis. More information about these treatments can be found in Methods and Appendices.}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\textbf{Table S1}. Perecentage of seeds produced per ovule for the ten
species used in the experiment. The treatments presented are hand
cross-pollination, hand self-pollination, natural selfing and apomixis
(emasculated flowers).
